---
title: "ESM 244 Lab 1"
author: "Tom Paschos"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives
- basic review
- intro to spatial data in R with sf
- intro to shiny apps 

1. Attach packages 

```{r packages}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)

```

2. Review: reading, exploring, wrangling and visualizing data

a. Read in the data

```{r}

wine_data <- read_csv("wine_data.csv")
View(wine_data)

wine <- read_csv("wine_data.csv")

```

b. wrangling

```{r}

# pipe operator = "and then" (Command + Shift + M)
# select = columns, filter = rows
# double == looks for a match, one = sets something new
# OR is vert line above return key

wine_new <- wine %>% 
  select(country, province, winery, region_1, points, price) %>% 
  rename(state = province) %>% 
  filter(state == "California" | state == "Oregon" | state == "Washington") %>% 
  mutate(ppd = points/price) %>% 
  arrange(-ppd)

```

c. Use group_by() and summarize () to find summary statistics by group

```{r}

# group_by can group by more than one variable
# This with summary is a great combo for data exploration

wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE) # na.rm = TRUE ignores NA values
  )

```


d. A few gg-graphics

```{r}

# ggplot needs 3 things: 1. say we're using ggplot 2. what our data is (and x and y variables), 3. what type of graph
# Piping and adding levels of a gg-graph are different things.  Use a + sign to add additional levels
# Looking at data separately by group, we can use facet wrap
# scales=free will make the scales different, which is harder to compare data with different scales, but that's okay

wine_hists <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
  facet_wrap(~state, scales = "free")

wine_hists

```

New Graph: scatterplot graph of price vs. points, with color determined by state, adding a linear model trendline on top of it.

```{r}

# If you don't reference a variable put it outside aes

wine_points <- ggplot(wine_new, aes(x=price, y=points)) +
  geom_point(aes(color = state), alpha = 0.5) +
  geom_smooth(color = "black", method = "lm", se = FALSE) + #adds a trendline
  scale_color_manual(values = c("royalblue", "mediumorchid1","orange")) +
  facet_wrap(~state) +
  theme_light()

wine_points

```



